### 2018-02-16 CO749-GraphColoring/assignments/A3.tex
finished a1

	Thus we have shown that $G$ is $k$-critical. It remains to verify that $G$ is splittable.

>>>>

	Thus we have shown that $G$ is $k$-critical. It remains to verify that $G$ is splittable. First consider $v \in V(G) \backslash\{z_x,z_y\}$. Consider $G'$, where we split $v$ into $v_1, \dots, v_m$ each with positive degree, $2 \leq m \leq k-1$. If $v \in V(G_1)$ then apply the corresponding split $v_1, \dots, v_m$ to $G_1$, and use that $G_1$ is splittable to obtain a $(k-1)$-coloring $\phi_1$. Observe $\phi_1(x) \neq \phi_1(y)$ since $xy \in E(G_1)$. Since $G_2$ is $2$-splittable, $G - G_1$ has a $(k-1)$-coloring $\phi_2$. Since $G_2$ is not $(k-1)$-colorable, $\phi_2(z_x) \neq \phi_2(z_y)$. Again we may permute colors of $\phi_2$ so $\phi_1$ and $\phi_2$ agree on $z_x$ and $z_y$, obtaining a $(k-1)$-coloring of $G'$.
	If $v \in V(G_2)$ then apply the corresponding split $v_1, \dots, v_m$ to $G_2$ and use splittability to obtain a $(k-1)$-coloring $\phi_2$. Setting $\phi_2(z_x) = \phi_2(z_y) = \phi_2(z)$ we obtain a $(k-1)$-coloring of $G-G_1$. By criticality, $G_1-xy$ has a $(k-1)$-coloring $\phi_1$. Since $G_1$ is not $(k-1)$-colorable, $\phi_1(x) = \phi_1(y)$. Hence we may permute colors of $\phi_2$ so $\phi_1$ and $\phi_2$ agree on $z_x$ and $z_y$, obtaining a $(k-1)$-coloring of $G'$.
	It remains to consider splitting one $z_x$ or $z_y$ in $G$. Without loss of generality let $G'$ be the graph obtained by splitting $z_x$ into $z_x^1, \dots, z_x^m$, each with positive degree, $2 \leq m \leq k-1$. Split $x$ in $G_1$ into $x^1, \dots, x^m$ with neighbourhoods $N(x^i) = N(z_x^i)\cap V(G_1)$ yielding graph $G_1'$. Obtain a $(k-1)$-coloring $\phi_1$ of $G_1'$ by splittability of $G$.
	\subparagraph{}Case $1:$ $\phi_1(x^i) = \phi_1(y)$ for some $i \in \{1,\dots, m\}$. Split $z$ in $G_2$ into $z^1,\dots, z^m$ with neighbourhoods $N(z^j) = N(z_x^j) \cap V(G_2)$ for $j \neq i$ and $$N(z^i) = (N(z_x^i) \cup N(z_y)) \cap V(G_2)$$
	obtaining graph $G_2'$. Color $G_2'$ by splittability of $G_2$ obtaining $(k-1)$-coloring $\phi_2$. Setting $\phi_2(z_y) = \phi_2(z^i)$ we obtain a $(k-1)$-coloring of $(G-G_1) \cup \{z_x^1, \dots, z_x^m, z_y\}$. Since $\phi_1$ and $\phi_2$ give all the split vertices different colors, and give $z_y$ the same color $z_x^i$ we may permute $\phi_1$ and $\phi_2$ to agree on $\{z_x^1, \dots, z_x^m, z_y\}$ obtaining a $(k-1)$-coloring of $G'$.
	\subparagraph{}Case $2:$ $\phi_1(x^i) \neq \phi_2(y)$ for all $i \in \{1,\dots, m\}$. Notice in this case $m < k-1$ by the pidgeonhole principle since $\phi_1$ assigns different colors to each split vertex. Split $z$ in $G_2$ into $z^1, \dots, z^m, z^{m+1}$ with neighbours $N(z^i) = N(z_x^i) \cap V(G_2)$ for $i \leq m$ and set $N(z^{m+1}) = N(z_y) \cap V(G_2)$. Applying splitability we obtain a $(k-1)$-coloring $\phi_2$. Now both $\phi_1$ and $\phi_2$ give $\{z_x^1, \dots, z_x^m, z_y\}$ different colors, so we may permute their colors to agree on $\{z_x^1, \dots, z_x^m, z_y\}$, and join the colorings to obtain a coloring of $G'$, as desired.
